---
title: "OptiVisT grasping analysis"
author: "Florian Pätzold, Ramon Zacharias"
date: "3/21/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Import all necessary packages.

```{r}
library(dplyr)
library(ggplot2)
library(EnvStats)
library(outliers)
```

# Data preprocessing

Set working directory, load all the grasping data, clean it and combine it into one dataframe.

```{r}
# Set working directory to the folder containing the CSV files
setwd("data")
# Get list of all CSV files for the grapsing task in the folder
file_list <- list.files(getwd(), pattern = "*grasping*")
# Delete testing data file
file_list <- file_list[file_list != "1111_grasping.csv"]
# Create an empty data frame to store the combined data
combined_data <- data.frame()
# Create inverted %in% function
`%ni%` <- Negate(`%in%`)
# Loop through each CSV file
for (file in file_list) {
  # Read the CSV file into a data frame
  file_data <- read.csv(file, header = TRUE, sep = ",")
  
  # Extract the number from the filename and add as a new column
  number <- as.numeric(gsub("[^0-9]+", "", file))
  file_data$participant_id <- number
  
  # More than 3 rep trials for each block is infeasible and must be wrong data saving
  # Then we only take the last 8 blocks
  if (length(file_data$location) > 8 * (9+3)) {
    file_data <- file_data[tail(which(file_data$time == "time"), n=1) + 1 : length(file_data$time), ]
    file_data <- file_data[complete.cases(file_data), ]
  }
  
  # Cast column types from factor to numeric/char
  file_data$time <- as.numeric(as.character(file_data$time))
  file_data$num_instructions <- as.numeric(as.character(file_data$num_instructions))
  file_data$location <- as.character(file_data$location)
  file_data$block <- as.numeric(as.character(file_data$block))
  
  # Assign the correct block number for repetition trials in the tactile condition of P 1-6
  transform(file_data, block = as.numeric(block))
  if (number <= 6 & file_data$condition[1] == "tactile") {
    for (i in 1:length(file_data$location)) {
      if (file_data$location[i] %ni% list(1,2,3,4,5,6,7,8,9,"location")) {
        file_data$block[i] <- file_data$block[i-1] # dtype of col is factor when it should be numeric
      }
    }
  }
  
  # Append the data to the combined data frame
  combined_data <- rbind(combined_data, file_data)
  
}

# Save the combined data frame as a CSV file
row.names(combined_data) <- NULL
write.csv(combined_data, "combined.csv")
```

# Data visualization

Calculate summary statistics and visualize aspects of the data.

```{r}
# Calculate the mean times for each participant
mean_times <- combined_data %>% 
  filter(success == "success") %>% 
  group_by(participant_id, condition) %>% 
  summarize(mean_time = mean(time))

# Save the mean trial times for each participant by condition
mean_times_auditory <- subset(mean_times, condition == "auditory")
mean_times_tactile <- subset(mean_times, condition == "tactile")
ggplot(mean_times, aes(x = condition, y = mean_time, fill = as.factor(participant_id))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Condition", y = "Mean time", fill = "Participant ID") +
  theme_bw()

# bar plot: x = fail, exfail (grouped by condition), y = counts
combined_data %>% group_by(condition, success) %>% count() %>% filter(success != "success") %>%
  ggplot(aes(x = condition, y = n, fill = success)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = n), position = position_dodge(width = 1), vjust = 2, hjust=0.5) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  labs(
    title = "Count of trials with false instruction or response", 
    x = NULL, 
    y = "Count \n",
    subtitle = "in auditory and tactile condition \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right", legend.title = element_blank()) + # legend.title = element_blank()
  scale_fill_discrete(labels = c("experimenter fail", "participant fail"))

# line plot: x = block (grouped by condition --> 2 lines), y = fail counts
combined_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail")) %>% 
  ggplot(aes(x = block, y = fail_count, color=condition)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Count of trials with false response per block by condition", 
    x = "\n Block number", 
    y = "Fail count \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(1:8)) +
  scale_y_continuous(breaks = c(0:8))

# density plot: x = median df$time (grouped by condition)
combined_data %>% group_by(condition) %>% filter(success == "success") %>% 
  ggplot(aes(x = time*1000, color=condition, fill=condition)) +
  geom_density(alpha=0.5) +
  labs(
    title = "Count of trials with false response per block by condition", 
    x = "\n Reaction time (ms)", 
    y = "\n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_y_continuous(breaks = c())

# box plot: x = condition, y = RT
combined_data %>% filter(success == "success") %>%
  ggplot(aes(x = condition, y = time*1000)) +
  geom_boxplot() +
  labs(
    title = "Mean and 25% and 75% reaction time quantiles per condition", 
    x = NULL, 
    y = "Reaction time (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank())

# line plot: x = block (grouped by condition --> 2 lines), y = median/mean RT
combined_data %>% group_by(block, condition) %>% summarize(mean_time = mean(time*1000)) %>% 
  ggplot(aes(x = block, y = mean_time, color=condition)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Mean reaction times per block by condition", 
    x = "\n Block number", 
    y = "Mean reaction times (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(1:8))

```

# Statistical analysis

### Normality assumption check

```{r}
# Split the data by condition
auditory_data <- combined_data[combined_data$condition == "auditory",]
tactile_data <- combined_data[combined_data$condition == "tactile",]
# Drop rows where success column is not equal to "success"
auditory_data <- auditory_data[auditory_data$success == "success",]
tactile_data <- tactile_data[tactile_data$success == "success",]
# Check for normality using histogram and normal probability plot for each condition
par(mfrow=c(2,2)) # create 2x2 plot grid
# Histogram and normal probability plot for auditory condition
hist(auditory_data$time, main="Histogram of Auditory Times", xlab="Times")
qqnorm(auditory_data$time, main="Normal Probability Plot of Auditory Times")
qqline(auditory_data$time)
# Histogram and normal probability plot for tactile condition
hist(tactile_data$time, main="Histogram of Tactile Times", xlab="Times")
qqnorm(tactile_data$time, main="Normal Probability Plot of Tactile Times")
qqline(tactile_data$time)
# Perform Shapiro-Wilk test for normality for each condition
shapiro.test(auditory_data$time)
shapiro.test(tactile_data$time)
# Perform Kolmogorov-Smirnov test for each condition
ks.test(auditory_data$time, "pnorm", mean(auditory_data$time), sd(auditory_data$time))
ks.test(tactile_data$time, "pnorm", mean(tactile_data$time), sd(tactile_data$time))
```

### Outlier detection

```{r}
# Grubbs' test by condition
grubbs_results <- by(combined_data$time, combined_data$condition, grubbs.test)
print(grubbs_results)
```

### Hypothesis testing

Perform the non-parametric Wilcoxon signed-rank test to compare group medians.

```{r}
# Perform Wilcoxon signed-rank test
wilcox.test(auditory_data$time, tactile_data$time, paired = FALSE)

# Perform Wilcoxon signed-rank test on the mean reaction times for each participant
wilcox.test(mean_times_auditory$mean_time, mean_times_tactile$mean_time, paired = TRUE)
```