---
title: "OptiVisT grasping analysis"
author: "Florian Pätzold, Ramon Zacharias"
date: "3/21/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Import all necessary packages.

```{r}
library(dplyr)
library(ggplot2)
library(EnvStats)
library(outliers)
library(lme4)
library(lmerTest) # significance testing for linear mixed models
```

# Data preprocessing

Set working directory, load all the grasping data, clean it and combine it into one dataframe.

```{r}
# Set working directory to the folder containing the CSV files
setwd("data")
# Get list of all CSV files for the grapsing task in the folder
file_list <- list.files(getwd(), pattern = "*grasping*")
# Delete testing data file
file_list <- file_list[file_list != "1111_grasping.csv"]
# Create an empty data frame to store the combined data
combined_data <- data.frame()
# Create inverted %in% function
`%ni%` <- Negate(`%in%`)
# Loop through each CSV file
for (file in file_list) {
  # Read the CSV file into a data frame
  file_data <- read.csv(file, header = TRUE, sep = ",")
  
  # Extract the number from the filename and add as a new column
  number <- as.numeric(gsub("[^0-9]+", "", file))
  file_data$participant_id <- number
  
  # More than 3 rep trials for each block is unfeasible and must be wrong data saving
  # Then we only take the last 8 blocks
  if (length(file_data$location) > 8 * (9+3)) {
    file_data <- file_data[tail(which(file_data$time == "time"), n=1) + 1 : length(file_data$time), ]
    file_data <- file_data[complete.cases(file_data), ]
  }
  
  # Cast column types from factor to numeric/char
  file_data$time <- as.numeric(as.character(file_data$time))
  file_data$num_instructions <- as.numeric(as.character(file_data$num_instructions))
  file_data$location <- as.character(file_data$location)
  file_data$block <- as.numeric(as.character(file_data$block))
  
  # Assign the correct block number for repetition trials in the tactile condition of P 1-6
  transform(file_data, block = as.numeric(block))
  if (number <= 6 & file_data$condition[1] == "tactile") {
    for (i in 1:length(file_data$location)) {
      if (file_data$location[i] %ni% list(1,2,3,4,5,6,7,8,9,"location")) {
        file_data$block[i] <- file_data$block[i-1] # dtype of col is factor when it should be numeric
      }
    }
  }
  
  # Append the data to the combined data frame
  combined_data <- rbind(combined_data, file_data)
  
}

# Save the combined data frame as a CSV file
row.names(combined_data) <- NULL
write.csv(combined_data, "combined.csv")
```

### Outlier detection

```{r}
# first block in each condition is training block
clean_data <- combined_data %>% filter(block != 1)

# Grubbs' test by condition
grubbs_results <- by(clean_data$time, clean_data$condition, grubbs.test)
print(grubbs_results)

# clean_data <- clean_data %>% filter(time < 10.0)
```

### Normality assumption check

```{r}
# Split the data by condition and drop fails
auditory_data <- clean_data %>% filter(condition == "auditory", success == "success")
tactile_data <- clean_data %>% filter(condition == "tactile", success == "success")

# Check for normality using histogram and normal probability plot for each condition
par(mfrow=c(2,2)) # create 2x2 plot grid

# Histogram and normal probability plot for auditory condition
hist(auditory_data$time, main="Histogram of Auditory Times", xlab="Times")
qqnorm(auditory_data$time, main="Normal Probability Plot of Auditory Times")
qqline(auditory_data$time)

# Histogram and normal probability plot for tactile condition
hist(tactile_data$time, main="Histogram of Tactile Times", xlab="Times")
qqnorm(tactile_data$time, main="Normal Probability Plot of Tactile Times")
qqline(tactile_data$time)

# Perform Shapiro-Wilk test for normality for each condition
shapiro.test(auditory_data$time)
shapiro.test(tactile_data$time)

# Perform Kolmogorov-Smirnov test for each condition (compares data to normal distribution pnorm)
ks.test(auditory_data$time, "pnorm", mean(auditory_data$time), sd(auditory_data$time))
ks.test(tactile_data$time, "pnorm", mean(tactile_data$time), sd(tactile_data$time))
```

The Shapiro-Wilk test (a,t: \*\*\*) suggests that the data was generated from a non-normal distribution --> we use non-parametric hypothesis tests --> Wilcoxon rank sum to compare unpaired data medians.

# Data visualiation and hypothesis testing

Calculate summary statistics, visualize aspects of the data and test the corresponding hypotheses.

```{r}
# Barplot: Mean trial times for each condition per participant
clean_data %>% filter(success == "success") %>% group_by(participant_id, condition) %>% summarize(mean_time = mean(time)*1000)
clean_data %>% filter(success == "success") %>% group_by(participant_id, condition) %>% summarize(mean_time = mean(time)*1000) %>% 
ggplot(aes(x = factor(participant_id), y = mean_time, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Mean trial times for each participant per condition", 
    x = "Participant", 
    y = "Trial time (ms) \n",
    fill = "condition"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right")
```

```{r}
# Bar plot: x = fail, exfail (grouped by condition), y = counts
clean_data %>% group_by(condition, success) %>% count() %>% filter(success != "success")
clean_data %>% group_by(condition, success) %>% count() %>% filter(success != "success") %>%
  ggplot(aes(x = condition, y = n, fill = success)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = n), position = position_dodge(width = 1), vjust = 2, hjust=0.5) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  labs(
    title = "Count of trials with false instruction or response", 
    x = NULL, 
    y = "Count \n",
    subtitle = "in auditory and tactile condition \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right", legend.title = element_blank()) +
  scale_fill_discrete(labels = c("experimenter fail", "participant fail"))
```

```{r}
# Line plot: x = block (grouped by condition --> 2 lines), y = fail counts
clean_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail"))
clean_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail")) %>% 
  ggplot(aes(x = block, y = fail_count, color=condition)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Count of trials with false response per block by condition", 
    x = "\n Block number", 
    y = "Fail count \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(1:8))
```

```{r}
# Density plot: x = time, y = probability density (grouped by condition)
clean_data %>% group_by(condition) %>% filter(success == "success") %>% 
  ggplot(aes(x = log(time*1000), color=condition, fill=condition)) +
  geom_density(alpha=0.5) +
  labs(
    title = "Density of log-transformed trial times by condition", 
    x = "\n Trial time (log(ms))", 
    y = "\n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank())
```

```{r}
# Box plot: x = condition, y = RT
clean_data %>% group_by(condition) %>% filter(success == "success") %>% summarize(mean_time = mean(time)*1000)
clean_data %>% filter(success == "success") %>%
  ggplot(aes(x = condition, y = time*1000)) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    title = "Mean and 25% and 75% trial time quantiles per condition", 
    x = NULL, 
    y = "Trial time (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_y_continuous(limits = c(0, 6000)) # remove outliers from plot
  

# Perform the non-parametric Wilcoxon rank sum (with unpaired data as fail trials are excluded) test to compare group medians.
# Compare the results to the results of the corresponding t-tests (group mean comparison for normally distributed data).
wilcox.test(auditory_data$time, tactile_data$time, paired = FALSE)
t.test(auditory_data$time, tactile_data$time, paired = FALSE)
```

```{r}
# Line plot: x = block (grouped by condition --> 2 lines), y = median/mean RT
combined_data %>% group_by(block, condition) %>% summarize(mean_time = mean(time)*1000)
combined_data %>% group_by(block, condition) %>% summarize(mean_time = mean(time)*1000) %>% 
  ggplot(aes(x = block, y = mean_time, color=condition)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Mean trial times per block by condition", 
    x = "\n Block number", 
    y = "Mean trial times (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(1:8))

# test diffs between blocks
# tactile
wilcox.test(filter(combined_data, block == 1, condition == "tactile")$time, 
            filter(combined_data, block == 2, condition == "tactile")$time, paired = FALSE) # ***
wilcox.test(filter(combined_data, block == 3, condition == "tactile")$time, 
            filter(combined_data, block == 4, condition == "tactile")$time, paired = FALSE) # *

# auditory
wilcox.test(filter(combined_data, block == 1, condition == "auditory")$time, 
            filter(combined_data, block == 2, condition == "auditory")$time, paired = FALSE) # ***

# other comparisons yielded no significant differences
```

```{r}
# Barplot: y = RT, x = fruit position
pos_data <- clean_data # dummy
pos_data$location[pos_data$location == "rep_1"] = "1"
pos_data$location[pos_data$location == "rep_2"] = "2"
pos_data$location[pos_data$location == "rep_3"] = "3"
pos_data$location[pos_data$location == "rep_4"] = "4"
pos_data$location[pos_data$location == "rep_5"] = "5"
pos_data$location[pos_data$location == "rep_6"] = "6"
pos_data$location[pos_data$location == "rep_7"] = "7"
pos_data$location[pos_data$location == "rep_8"] = "8"
pos_data$location[pos_data$location == "rep_9"] = "9"

pos_data %>% filter(success == "success") %>% group_by(condition, location) %>% summarize(mean_time = mean(time)*1000)
pos_data %>% filter(success == "success") %>% group_by(condition, location) %>% summarize(mean_time = mean(time)*1000) %>%
  ggplot(aes(x = location, y = mean_time, fill = condition)) +
  geom_bar(stat="identity", position = "dodge") +
  labs(
    title = "Mean trial time per fruit location by condition", 
    x = "\n Fruit position", 
    y = "Mean trial time (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right")


# mid (5) fastest as starting point is in front
# one-command positions up (2), left (4), right (6), down (8) are shorter than two-command positions
# diff between conditions for positions 1 and 3 (upper left and right) greater than for 7 and 9 (lower left and right)
# --> grasping at top shelf took longer, probably the chair was too low

# Could add test for each fruit position comparing conditions
# Could add comparison between upper left (1) and upper right (3) and/or lower left (7) and lower right (9)
```

```{r}
# Subset data by order
data_tfirst <- clean_data[clean_data$participant_id == 2 
                             | clean_data$participant_id == 3 
                             | clean_data$participant_id == 5 
                             | clean_data$participant_id == 8 
                             | clean_data$participant_id == 10 
                             | clean_data$participant_id == 12 
                             | clean_data$participant_id == 15 
                             | clean_data$participant_id == 17 
                             | clean_data$participant_id == 19 
                             | clean_data$participant_id == 20
                             | clean_data$participant_id == 21 
                             | clean_data$participant_id == 24 
                             | clean_data$participant_id == 26 
                             | clean_data$participant_id == 27 
                             | clean_data$participant_id == 29, ]
data_tfirst$order <- "tactile_auditory"

data_afirst <- clean_data[clean_data$participant_id == 1 
                             | clean_data$participant_id == 4 
                             | clean_data$participant_id == 6 
                             | clean_data$participant_id == 7 
                             | clean_data$participant_id == 9 
                             | clean_data$participant_id == 11 
                             | clean_data$participant_id == 13 
                             | clean_data$participant_id == 14 
                             | clean_data$participant_id == 16 
                             | clean_data$participant_id == 18
                             | clean_data$participant_id == 22 
                             | clean_data$participant_id == 23 
                             | clean_data$participant_id == 25 
                             | clean_data$participant_id == 28 
                             | clean_data$participant_id == 30, ]
data_afirst$order <- "auditory_tactile"

# bind the order data
order_data <- rbind(data_tfirst, data_afirst)

# Boxplot: x = condition, y = trial time, grouped by order
order_data %>% filter (success == "success") %>% group_by(condition, order) %>% summarize(mean_time = mean(time)*1000)
order_data %>% filter (success == "success") %>% 
  ggplot(aes(x = condition, y = time*1000, fill = order)) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    title = "Mean trial times per condition by condition order", 
    x = NULL, 
    y = "Trial time (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right") +
  scale_fill_discrete(labels = c("auditory first", "tactile first")) +
  scale_y_continuous(limits = c(0, 6000)) # remove outliers from plot

# test diffs for significance
tactile_tfirst <- order_data %>% filter(success == "success", condition == "tactile", order == "tactile_auditory")
tactile_afirst <- order_data %>% filter(success == "success", condition == "tactile", order == "auditory_tactile")
auditory_tfirst <- order_data %>% filter(success == "success", condition == "auditory", order == "tactile_auditory")
auditory_afirst <- order_data %>% filter(success == "success", condition == "auditory", order == "auditory_tactile")
wilcox.test(tactile_tfirst$time, tactile_afirst$time, paired = FALSE) # ***
wilcox.test(auditory_tfirst$time, auditory_afirst$time, paired = FALSE) # **
```

```{r}
# Subset data by experimenter
data_e1 <- clean_data[clean_data$participant_id == 1 
                         | clean_data$participant_id == 3 
                         | clean_data$participant_id == 14 
                         | clean_data$participant_id == 17 
                         | clean_data$participant_id == 18 
                         | clean_data$participant_id == 20
                         | clean_data$participant_id == 28
                         | clean_data$participant_id == 29
                         | clean_data$participant_id == 30, ]
data_e1$experimenter <- "1" # m

data_e2 <- clean_data[clean_data$participant_id == 2 
                         | clean_data$participant_id == 4 
                         | clean_data$participant_id == 5 
                         | clean_data$participant_id == 6 
                         | clean_data$participant_id == 10 
                         | clean_data$participant_id == 11 
                         | clean_data$participant_id == 12 
                         | clean_data$participant_id == 13
                         | clean_data$participant_id == 21
                         | clean_data$participant_id == 22
                         | clean_data$participant_id == 26, ]
data_e2$experimenter <- "2" # p

data_e3 <- clean_data[clean_data$participant_id == 7 
                         | clean_data$participant_id == 8 
                         | clean_data$participant_id == 9 
                         | clean_data$participant_id == 15 
                         | clean_data$participant_id == 16 
                         | clean_data$participant_id == 19
                         | clean_data$participant_id == 23
                         | clean_data$participant_id == 24
                         | clean_data$participant_id == 25
                         | clean_data$participant_id == 27, ]
data_e3$experimenter <- "3" # f

# bind the data together
experimenter_data <- rbind(data_e1, data_e2, data_e3)

# Boxplot: x = experimenter, y = trial time
experimenter_data %>% filter(success == "success") %>% group_by(experimenter) %>% summarize(mean_times = mean(time)*1000)
experimenter_data %>% filter(success == "success") %>% 
  ggplot(aes(x = experimenter, y = time*1000)) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    title = "Mean trial times per experimenter", 
    x = "Experimenter", 
    y = "Trial time (ms) \n",
    color = "experimenter"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right") +
  scale_fill_discrete(labels = c("auditory first", "tactile first")) +
  scale_y_continuous(limits = c(0, 6000)) # remove outliers from plot

# Barplot: x = experimenter, y = number of fails, grouped by type of fail
experimenter_data %>% group_by(experimenter, success) %>% count() %>% filter(success != "success")
experimenter_data %>% group_by(experimenter, success) %>% count() %>% filter(success != "success") %>%
  ggplot(aes(x = experimenter, y = n, fill = success)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = n), position = position_dodge(width = 1), vjust = 2, hjust=0.5) +
  labs(
    title = "Count of failed trials per experimenter by failure type", 
    x = "Experimenter", 
    y = "Count \n",
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right", legend.title = element_blank()) + # legend.title = element_blank()
  scale_fill_discrete(labels = c("experimenter fail", "participant fail"))
```
# Bind data for LMM

```{r}
grasping_data <- clean_data %>% filter(success == "success") %>% select(-success)

# add experimenter to df
grasping_data$experimenter <- ifelse(grasping_data$participant_id %in% c(1,3,14,17,18,20,28,29,30), "1", # m
                           ifelse(grasping_data$participant_id %in% c(2,4,5,6,10,11,12,13,21,22,26), "2", "3")) # p, f

# add condition to df
grasping_data$order <- ifelse(grasping_data$participant_id %in% c(2,3,5,8,10,12,15,17,19,20,21,24,26,27,29),
                              "tactile_first", "auditory_first")
grasping_data$order <- ifelse(grasping_data$participant_id %in% c(2,3,5,8,10,12,15,17,19,20,21,24,26,27,29),
                              "tactile_first", "auditory_first")

# rename repetition locations in df
grasping_data$location[grasping_data$location == "rep_1"] = "1"
grasping_data$location[grasping_data$location == "rep_2"] = "2"
grasping_data$location[grasping_data$location == "rep_3"] = "3"
grasping_data$location[grasping_data$location == "rep_4"] = "4"
grasping_data$location[grasping_data$location == "rep_5"] = "5"
grasping_data$location[grasping_data$location == "rep_6"] = "6"
grasping_data$location[grasping_data$location == "rep_7"] = "7"
grasping_data$location[grasping_data$location == "rep_8"] = "8"
grasping_data$location[grasping_data$location == "rep_9"] = "9"
grasping_data$time <- grasping_data$time*1000 # s -> ms

# cast block number and number of instructions to characters
grasping_data$block <- as.character(grasping_data$block) # numeric to character for linear mixed model
grasping_data$num_instructions <- as.character(grasping_data$num_instructions)
```

# Exploration

In order to decide which variables are included in the LMM and which are excluded, we evaluate models, compare them, test for significant effects and investigate their complexity in terms of number of coefficients/parameters.

- LMM Tutorial: https://www.youtube.com/watch?v=QCqF-2E86r0 !!!
- Interpretation explanation: https://www.youtube.com/watch?v=yJnHmCMb1q4
- An LMM example (interaction explanation): https://www.youtube.com/watch?v=W8txfclM16U

Decision train of thought:
- variables: num_instructions, location, block, condition, pID, order, experimenter
- fixed effects we are interested in: condition (difference between sensory modalities), block (learning effects)
- num_instructions and location correlate (num is inferable from loc) and location gives a little more exact information (not only instructions = 2, but also whether top-left, top-right or bottom-left, bottom-right) but makes the model drastically more complex --> we would choose num_instructions as fixed --> we are not interested in it so we exclude
- crossed random effects: order and experimenter as some participants did have the order tactile_first, but not the other order, or did the experiment with "f" as first experimenter and not with the other, i.e. random effect 'order' occurs in each of the levels of random effect 'experimenter'
- we also model varying condition slopes of both random effects but not block slopes as the blocks in principle have the same structure, while the conditions are inherently different
- usually random effects should have at least 5 levels, but I want to include them  
- in theory participantID could also be seen as random effect, but I think this is the variation we want to investigate in science -> check for interaction effects (correlations?) between variables and leave out variables without significant effect 
- (model comparison: anova with model h0 with additional variable, h1 without additional var) --> can be replaced by lmerTest package anova testing

```{r}
# model testing: decide which variables to include/exclude (with lmerTest)
anova(model_complex <- lmer(time ~ condition * block * (condition | order) + (condition | experimenter), 
                            data = grasping_data))
anova(model_complex <- lmer(time ~ condition * block * location * (condition | order) + (condition | experimenter), 
                            data = grasping_data))
# --> introduction of location makes interaction effect significant, but we are not interested in location, so we do not investigate the interaction effects either (not significant AND not interested)

# linear mixed model: complex (with significant interactions)
model_complex <- lmer(time ~
                condition * # include or exclude interaction?
                block +
                location +
                condition:location + # include or exclude interaction?
                # show correlated varying condition slopes and intercepts for each random effect
                (condition | order) +
                (condition | experimenter), 
              data = grasping_data)

anova(model_complex) # with lmerTest package
summary(model_complex)
coef(model_complex)

# linear mixed model: simple (without interactions)
model_occam <- lmer(time ~
                condition +
                block +
                location +
                (condition | order) +
                (condition | experimenter), 
              data = grasping_data)

anova(model_occam)
summary(model_occam)
coef(model_occam)

# alternative (simpler) model: 'number of instructions' instead 'location' for less parameters
model_alt1 <- lmer(time ~
                condition +
                block +
                num_instructions +
                (condition | order) +
                (condition | experimenter), 
              data = grasping_data)
summary(model_alt1)

# alternative (simplest) model: disregard 'block' entirely
model_alt2 <- lmer(time ~
                condition +
                num_instructions +
                (condition | order) +
                (condition | experimenter), 
              data = grasping_data)
summary(model_alt2)
```
# Final LMM

```{r}
model <- lmer(time ~
                condition +
                block +
                (condition | order) +
                (condition | experimenter), 
              data = grasping_data)

# anova(model) # --> no significant difference in trial times between conditions in this model
# coef(model) #
summary(model) 
# "boundary (singular) fit: see ?isSingular" --> model too complex, cannot converge -> leave out block?
# without 'block' it gives the same warning --> include
```

Random effects cannot be plotted against time since experimenter, order (and block) are all discrete.

# Model diagnostics

```{r}
# 1. plot residuals vs. fitted for entire model & vs. each independent var -> should show no trends, normally distributed, equal variance

# 2. plot residuals vs. fitted for each random effect --> should show no trends, normally distributed, equal variance

# 3. conditional mean of random effects should be normally distributed (normal QQ-plot)
```


