---
title: "OptiVisT grasping task analysis (Blind)"
author: "Florian Pätzold, Ramon Zacharias"
date: "3/21/2023"
output:
  html_document: default
  pdf_document: default
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Import all necessary packages.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(sjPlot)
library(gridExtra)
library(EnvStats)
library(outliers)
library(lme4)
library(lmerTest) # significance testing for linear mixed models
```

# Data preprocessing

Set working directory, load all the grasping data, clean it and combine it into one data frame.

```{r}
# Set working directory to the folder containing the CSV files
SAVE <- paste0(getwd(), "/Plots/")
setwd("../Data/Blind")
# Get list of all CSV files for the grapsing task in the folder
file_list <- list.files(getwd(), pattern = "*grasping*")
# Delete testing data file
file_list <- file_list[file_list != "1111_grasping.csv"]
# Create an empty data frame to store the combined data
combined_data <- data.frame()
# Create inverted %in% function
`%ni%` <- Negate(`%in%`)
# Loop through each CSV file
for (file in file_list) {
  # Read the CSV file into a data frame
  file_data <- read.csv(file, header = TRUE, sep = ",")
  
  # Extract the number from the filename and add as a new column
  number <- as.numeric(gsub("[^0-9]+", "", file))
  file_data$participant_id <- number
  
  # More than 3 rep trials for each block is unfeasible and must be wrong data saving
  # Then we only take the last 8 blocks
  if (length(file_data$location) > 8 * (9+3)) {
    file_data <- file_data[tail(which(file_data$time == "time"), n=1) + 1 : length(file_data$time), ]
    file_data <- file_data[complete.cases(file_data), ]
  }
  
  # Cast column types from factor to numeric/char
  file_data$time <- as.numeric(as.character(file_data$time))
  file_data$num_instructions <- as.numeric(as.character(file_data$num_instructions))
  file_data$location <- as.character(file_data$location)
  file_data$block <- as.numeric(as.character(file_data$block))
  
  # Assign the correct block number for repetition trials in the tactile condition of P 1-6
  transform(file_data, block = as.numeric(block))
  if (number <= 6 & file_data$condition[1] == "tactile") {
    for (i in 1:length(file_data$location)) {
      if (file_data$location[i] %ni% list(1,2,3,4,5,6,7,8,9,"location")) {
        file_data$block[i] <- file_data$block[i-1] # dtype of col is factor when it should be numeric
      }
    }
  }
  
  # Append the data to the combined data frame
  combined_data <- rbind(combined_data, file_data)
}

# Save the combined data frame as a CSV file
row.names(combined_data) <- NULL
write.csv(combined_data, "combined_blind.csv")
blindfolded <- read.csv("../combined.csv") %>% filter(block != 1)
```

## Outlier detection

```{r}
# first block in each condition is training block
clean_data <- combined_data %>% filter(block != 1)
clean_data$num_instructions <- ifelse(clean_data$location %in% c(5), 1, 
                                      ifelse(clean_data$location %in% c(2, 4, 6, 8), 2, 3))

# Grubbs' test by condition
grubbs_results <- by(clean_data$time, clean_data$condition, grubbs.test)
print(grubbs_results)

# clean_data <- clean_data %>% filter(time < 10.0)
```

## Normality assumption check

```{r}
# Split the data by condition and drop fails
auditory_data <- clean_data %>% filter(condition == "auditory", success == "success")
tactile_data <- clean_data %>% filter(condition == "tactile", success == "success")

# Check for normality using histogram and normal probability plot for each condition
par(mfrow=c(2,2)) # create 2x2 plot grid

# Histogram and normal probability plot for auditory condition
hist(auditory_data$time, main="Histogram of Auditory Times", xlab="Times")
qqnorm(auditory_data$time, main="Normal Probability Plot of Auditory Times")
qqline(auditory_data$time)

# Histogram and normal probability plot for tactile condition
hist(tactile_data$time, main="Histogram of Tactile Times", xlab="Times")
qqnorm(tactile_data$time, main="Normal Probability Plot of Tactile Times")
qqline(tactile_data$time)

# normal distribution
x <- seq(min(log(tactile_data$time)), max(log(tactile_data$time)), length.out = length(tactile_data$time))
y <- dnorm(x, mean = mean(log(tactile_data$time)), sd = sd(log(tactile_data$time)))
t_normal <- data.frame(x = x, y = y)

x <- seq(min(log(auditory_data$time)), max(log(auditory_data$time)), length.out = length(auditory_data$time))
y <- dnorm(x, mean = mean(log(auditory_data$time)), sd = sd(log(auditory_data$time)))
a_normal <- data.frame(x = x, y = y)

# plot density comparisons for t and a (justification for parametric LMM)
ggplot(tactile_data) + 
  geom_density(aes(x = log(time)), fill = "red", alpha = 0.5) +
  geom_line(aes(x = t_normal$x, y = t_normal$y), size = 0.7, color = "darkgreen")

ggplot(auditory_data) + 
  geom_density(aes(x = log(time)), fill = "red", alpha = 0.5) +
  geom_line(aes(x = a_normal$x, y = a_normal$y), size = 0.7, color = "darkgreen")
```
```{r, eval=FALSE, include=FALSE}
# Perform Shapiro-Wilk test for normality for each condition
shapiro.test(auditory_data$time)
shapiro.test(tactile_data$time)

# Perform Kolmogorov-Smirnov test for each condition (compares data to normal distribution pnorm)
ks.test(auditory_data$time, "pnorm", mean(auditory_data$time), sd(auditory_data$time))
ks.test(tactile_data$time, "pnorm", mean(tactile_data$time), sd(tactile_data$time))
```

Looks multi-modal because of different number of instructions.

```{r}
# Assumption: tri-modality of the data for different number of commands
aud_1 <- clean_data %>% filter(condition == "auditory", success == "success", num_instructions == 1)
aud_2 <- clean_data %>% filter(condition == "auditory", success == "success", num_instructions == 2)
aud_3 <- clean_data %>% filter(condition == "auditory", success == "success", num_instructions == 3)

tac_1 <- clean_data %>% filter(condition == "tactile", success == "success", num_instructions == 1)
tac_2 <- clean_data %>% filter(condition == "tactile", success == "success", num_instructions == 2)
tac_3 <- clean_data %>% filter(condition == "tactile", success == "success", num_instructions == 3)

# plot density comparisons filtered by number of commands
compare_to_normal <- function(data) {
  x <- seq(min(data$time), max(data$time), length.out = length(data$time))
  y <- dnorm(x, mean = mean(data$time), sd = sd(data$time))
  t_normal <- data.frame(x = x, y = y)
  
  p <- ggplot(data) + 
    geom_density(aes(x = time), fill = "red", alpha = 0.5) +
    geom_line(aes(x = t_normal$x, y = t_normal$y), size = 0.7, color = "darkgreen") +
    ggtitle(sprintf("data: %s", deparse(substitute(data))))
  
  return(p)
}

aud1 <- compare_to_normal(aud_1)
aud2 <- compare_to_normal(aud_2)
aud3 <- compare_to_normal(aud_3)

tac1 <- compare_to_normal(tac_1)
tac2 <- compare_to_normal(tac_2)
tac3 <- compare_to_normal(tac_3)

grid.arrange(aud1,aud2,aud3,tac1,tac2,tac3, ncol = 3)

#ggsave(paste0(SAVE,"normality_blind.jpeg"), plot = grid, dpi = 600)
```
```{r, eval=FALSE, include=FALSE}
# Statistical test
shapiro.test(log(aud_1$time))
shapiro.test(log(aud_2$time))
shapiro.test(log(aud_3$time))

shapiro.test(log(tac_1$time))
shapiro.test(log(tac_2$time))
shapiro.test(log(tac_3$time))
```

Statistical tests suggest non-normality, visually they are normally distributed. The t-test and Wilcoxon rank sum test yield the same results in all comparisons, so we go with the student's t-test and assume that the data stems from a normal distribution.

# Data visualisation and hypothesis testing

Calculate summary statistics, visualize aspects of the data and test the corresponding hypotheses.

## Trial times per participant

```{r}
# time standard devs
clean_data %>% group_by(condition) %>% summarize(mean = mean(time), sd = sd(time))
combined_data %>% group_by(condition, block) %>% summarize(mean = mean(time), sd = sd(time))
```


```{r}
times_per_participant <- clean_data %>% 
  filter(success == "success") %>% 
  group_by(participant_id, condition) %>% 
  summarize(mean_time = mean(time) * 1000) %>% 
  spread(condition, mean_time) %>% 
  mutate(diff = tactile - auditory, color = ifelse(diff > 0, "#00BFC4", "#F8766D")) %>% 
  ggplot(aes(x = factor(participant_id), y = diff, fill = color)) +
  geom_bar(stat = "identity", position = "identity") +
  labs(
    title = "Mean trial time difference between conditions per participant",
    x = "\n Participant",
    y = "Trial time difference (ms) \n") +
  theme_linedraw() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right") +
  scale_fill_manual(name=NULL, values = c("#00BFC4", "#F8766D"), labels = c("Tactile > Auditory", "Auditory > Tactile"))

times_per_participant
#ggsave(paste0(SAVE,"times_diff_per_participant_blind.jpeg"), plot = times_per_participant, dpi = 600)
```

```{r}
# Barplot: Mean trial times for each condition per participant
clean_data %>% filter(success == "success") %>% group_by(participant_id, condition) %>% summarize(mean_time = mean(time)*1000)
times_per_participant <- clean_data %>% 
  filter(success == "success") %>% 
  group_by(participant_id, condition) %>% 
  summarize(mean_time = mean(time)*1000) %>% 
  ggplot(aes(x = factor(participant_id), y = mean_time, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Mean trial times for each participant per condition", 
    x = "Participant", 
    y = "Trial time (ms) \n",
    fill = "condition"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right")

times_per_participant
#ggsave(paste0(SAVE,"times_per_participant_blind.jpeg"), plot = times_per_participant, dpi = 600)
```

## Trial times per condition

```{r}
blindfolded_violin <- blindfolded %>% filter(success == "success") %>% select(-1)
blindfolded_violin$blindedness <- "Blindfolded"
blindfolded_violin %>% summarize(mean_time = mean(time)*1000)
blind_violin <- clean_data %>% filter(success == "success")
blind_violin$blindedness <- "Blind"
blind_violin %>% summarize(mean_time = mean(time)*1000)

violins <- rbind(blindfolded_violin, blind_violin)

# Violin plot: x = condition, y = RT
times_condition_violin <- violins %>%
  ggplot(aes(x = blindedness, y = time*1000, color=condition)) +
  geom_violin(width=0.7) +
  geom_boxplot(outlier.shape = NA, width=0.2, position = position_dodge(width = 0.7)) +
  labs(
    #title = "Distribution of trial times per condition", 
    x = NULL, 
    y = "Trial time (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) #+
  #scale_y_continuous(limits = c(0, 6000)) # remove outliers from plot

times_condition_violin
#ggsave(paste0(SAVE,"times_condition_violin_blind.jpeg"), plot = times_condition_violin, dpi = 600)


# Perform t-test (with unpaired data as fail trials are excluded) test to compare group means
t.test(auditory_data$time, tactile_data$time, paired = FALSE)
```

```{r}
learning_data <- combined_data %>% filter(success == "success")

# Line plot: x = block (grouped by condition --> 2 lines), y = median/mean RT
learning_data %>% group_by(block, condition) %>% summarize(mean_time = mean(time)*1000)
times_per_block <- learning_data %>% group_by(block, condition) %>% summarize(mean_time = mean(time)*1000) %>% 
  ggplot(aes(x = block, y = mean_time, color=condition)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Mean trial times per block by condition", 
    x = "\n Block number", 
    y = "Mean trial times (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(1:8)) #+
  #scale_y_continuous(limits = c(2000, 6000))

times_per_block
#ggsave(paste0(SAVE,"times_per_block_blind.jpeg"), plot = times_per_block, dpi = 600)

# test diffs between blocks
# tactile
t.test(filter(learning_data, block == 1, condition == "tactile")$time, 
            filter(learning_data, block == 2, condition == "tactile")$time, paired = FALSE) # ***
t.test(filter(learning_data, block == 3, condition == "tactile")$time, 
            filter(learning_data, block == 4, condition == "tactile")$time, paired = FALSE) # *

# auditory
t.test(filter(learning_data, block == 1, condition == "auditory")$time, 
            filter(learning_data, block == 2, condition == "auditory")$time, paired = FALSE) # ***

# other comparisons yielded no significant differences
t.test(filter(learning_data, block == 4, condition == "auditory")$time, 
            filter(learning_data, block == 5, condition == "auditory")$time, paired = FALSE)

t.test(filter(learning_data, block == 4, condition == "tactile")$time, 
            filter(learning_data, block == 5, condition == "tactile")$time, paired = FALSE)
```

```{r}
# Barplot: y = RT, x = fruit position
pos_data <- clean_data # dummy
pos_data$location[pos_data$location == "rep_1"] = "1"
pos_data$location[pos_data$location == "rep_2"] = "2"
pos_data$location[pos_data$location == "rep_3"] = "3"
pos_data$location[pos_data$location == "rep_4"] = "4"
pos_data$location[pos_data$location == "rep_5"] = "5"
pos_data$location[pos_data$location == "rep_6"] = "6"
pos_data$location[pos_data$location == "rep_7"] = "7"
pos_data$location[pos_data$location == "rep_8"] = "8"
pos_data$location[pos_data$location == "rep_9"] = "9"

pos_data %>% filter(success == "success") %>% group_by(condition, location) %>% summarize(mean_time = mean(time)*1000)
pos_data %>% filter(success == "success") %>% group_by(condition, location) %>% summarize(mean_time = mean(time)*1000) %>%
  ggplot(aes(x = location, y = mean_time, fill = condition)) +
  geom_bar(stat="identity", position = "dodge") +
  labs(
    title = "Mean trial time per fruit location by condition", 
    x = "\n Fruit position", 
    y = "Mean trial time (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right")


# mid (5) fastest as starting point is in front
# one-command positions up (2), left (4), right (6), down (8) are shorter than two-command positions
# diff between conditions for positions 1 and 3 (upper left and right) greater than for 7 and 9 (lower left and right)
# --> grasping at top shelf took longer, probably the chair was too low

# Could add test for each fruit position comparing conditions
# Could add comparison between upper left (1) and upper right (3) and/or lower left (7) and lower right (9)
```

```{r}
# Subset data by order
data_afirst <- clean_data[clean_data$participant_id == 777 
                             | clean_data$participant_id == 778, ]


if (length(data_afirst != 0)) {
  data_afirst$order <- "auditory_tactile"
}

# Boxplot: x = condition, y = trial time, grouped by order
data_afirst %>% filter (success == "success") %>% group_by(condition, order) %>% summarize(mean_time = mean(time)*1000)
data_afirst %>% filter (success == "success") %>% 
  ggplot(aes(x = condition, y = time*1000, fill = order)) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    title = "Mean trial times per condition by condition order", 
    x = NULL, 
    y = "Trial time (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right") +
  scale_fill_discrete(labels = c("auditory first", "tactile first")) +
  scale_y_continuous(limits = c(0, 6000)) # remove outliers from plot
```

## Fails by condition

```{r}
# Bar plot: x = fail, exfail (grouped by condition), y = counts
clean_data %>% group_by(condition, success) %>% count() %>% filter(success != "success")
clean_data %>% group_by(condition, success) %>% count() %>% filter(success != "success") %>%
  ggplot(aes(x = condition, y = n, fill = success)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = n), position = position_dodge(width = 1), vjust = 2, hjust=0.5) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  labs(
    title = "Count of trials with false instruction or response", 
    x = NULL, 
    y = "Count \n",
    subtitle = "in auditory and tactile condition \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right", legend.title = element_blank()) +
  scale_fill_discrete(labels = c("experimenter fail", "participant fail"))
```

```{r}
# Line plot: x = block (grouped by condition --> 2 lines), y = fail counts
clean_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail"), exFail_count = sum(success == "exFail"))
fails_per_block <- clean_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail"), exFail_count = sum(success == "exFail")) %>% 
  ggplot(aes(x = block, y = fail_count, color=condition)) +
  geom_point(aes(y = fail_count, shape="Group 1"), size=3) +
  geom_line(aes(y = fail_count)) +
  geom_point(aes(y = exFail_count, shape="Group 2"), size=3) +
  geom_line(aes(y = exFail_count)) +
  labs(
    title = "Count of trials with false response or instruction per block by condition", 
    x = "\n Block number", 
    y = "Fail count \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(1:8)) +
  scale_shape_manual(
    name = "Fail type",
    values = c("Group 1" = 16, "Group 2" = 5),  # Use 16 for a solid point
    labels = c("Participant fails", "Experimenter fails"))

fails_per_block
#ggsave(paste0(SAVE,"fails_per_block_blind.jpeg"), plot = fails_per_block, dpi = 600)

# test for significance
fails_a <- clean_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail")) %>% filter(condition=="auditory")
fails_t <- clean_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail")) %>% filter(condition=="tactile")

t.test(fails_a$fail_count, fails_t$fail_count)

fails_time <- clean_data %>% group_by(block, condition) %>% summarize(fail_count = sum(success == "fail"), mean_time = mean(time)) %>% filter(condition=="tactile")
cor.test(fails_time$mean_time, fails_time$fail_count)
```


## Fails per experimenter

```{r}
# Subset data by experimenter
data_e2 <- clean_data[clean_data$participant_id == 778, ]
data_e2$experimenter <- "2" # p

data_e3 <- clean_data[clean_data$participant_id == 777, ]
data_e3$experimenter <- "3" # f

# bind the data together
experimenter_data <- rbind(data_e2, data_e3)

# Boxplot: x = experimenter, y = trial time
experimenter_data %>% filter(success == "success") %>% group_by(experimenter) %>% summarize(mean_times = mean(time)*1000)
experimenter_data %>% filter(success == "success") %>% 
  ggplot(aes(x = experimenter, y = time*1000)) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    title = "Mean trial times per experimenter", 
    x = "Experimenter", 
    y = "Trial time (ms) \n",
    color = "experimenter"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right") +
  scale_fill_discrete(labels = c("auditory first", "tactile first")) +
  scale_y_continuous(limits = c(0, 6000)) # remove outliers from plot

# Barplot: x = experimenter, y = number of fails, grouped by type of fail
experimenter_data %>% group_by(experimenter, success) %>% count() %>% filter(success != "success")
experimenter_data %>% group_by(experimenter, success) %>% count() %>% filter(success != "success") %>%
  ggplot(aes(x = experimenter, y = n, fill = success)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = n), position = position_dodge(width = 1), vjust = 2, hjust=0.5) +
  labs(
    title = "Count of failed trials per experimenter by failure type", 
    x = "Experimenter", 
    y = "Count \n",
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right", legend.title = element_blank()) + # legend.title = element_blank()
  scale_fill_discrete(labels = c("experimenter fail", "participant fail"))
```

# Linear Mixed-Effects Model (LMM)

## Data preparation

```{r}
grasping_data <- clean_data %>% filter(success == "success") %>% select(-success)

# add experimenter to df
grasping_data$experimenter <- ifelse(grasping_data$participant_id %in% c(778), "2", "3") # p, f

# add condition to df
grasping_data$order <- ifelse(grasping_data$participant_id %in% c(777,778), "auditory_first", "tactile_first")

# rename repetition locations in df
grasping_data$location[grasping_data$location == "rep_1"] = "1"
grasping_data$location[grasping_data$location == "rep_2"] = "2"
grasping_data$location[grasping_data$location == "rep_3"] = "3"
grasping_data$location[grasping_data$location == "rep_4"] = "4"
grasping_data$location[grasping_data$location == "rep_5"] = "5"
grasping_data$location[grasping_data$location == "rep_6"] = "6"
grasping_data$location[grasping_data$location == "rep_7"] = "7"
grasping_data$location[grasping_data$location == "rep_8"] = "8"
grasping_data$location[grasping_data$location == "rep_9"] = "9"
grasping_data$time <- grasping_data$time*1000 # s -> ms

# cast block number and number of instructions to characters
grasping_data$block <- as.character(grasping_data$block) # numeric to character for linear mixed model
grasping_data$num_instructions <- as.character(grasping_data$num_instructions)
```

## Final LMM

```{r}
model <- lmer(time ~
                condition +
                block +
                (condition | experimenter), 
              data = grasping_data)

summary(model)
#anova(model) # --> no significant difference in trial times between conditions in this model
```
Random effects cannot be plotted against time since experimenter, order (and block) are all discrete.

### LMM Visualizations

```{r}
grasping_data$block <- as.double(grasping_data$block)

# change label names
labels <- c(`1` = "Experimenter 1",
            `2` = "Experimenter 2",
            `3` = "Experimenter 3",
            `auditory_first` = "Auditory First",
            `tactile_first` = "Tactile First")

lmm_plot <- grasping_data %>% group_by(block, condition, experimenter, order) %>% summarize(mean_time = mean(time)) %>% 
  ggplot(aes(x = block, y = mean_time, color=condition)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Mean trial times per block by condition", 
    x = "\n Block number", 
    y = "Mean trial times (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(2:8)) +
  #scale_y_continuous(limits = c(2000, 4000)) +
  facet_grid(experimenter ~ order, 
             labeller = as_labeller(labels))
lmm_plot

#ggsave(paste0(SAVE,"lmm_plot_blind.jpeg"), plot = lmm_plot, dpi = 600)
```



