---
title: "OptiVisT localization analysis"
author: "Florian Pätzold, Ramon Zacharias"
date: "2023-07-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Import all necessary packages.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(plotly)
```

# Data preprocessing

Set working directory, load all the localization data, clean it and combine it into one data frame. Also load the grasping data for later comparisons.

```{r}
# Set working directory to the folder containing the CSV files
HOME <- getwd()
setwd(paste0(HOME, "/data"))

# Load combined grasping data
grasping_data <- read.csv("combined.csv", header = TRUE, sep = ",")
#grasping_data
```

```{r}
# Create a list of file names that end with "_localization.csv"
setwd(paste0(HOME, "/Data"))
file_list <- list.files(pattern = "_localization*")
# Create an empty data frame to store the combined data
localization_data <- data.frame()
# loop through each file in the list and add the data to combined_data
for (file in file_list) {
  # Extract the participant ID from the file name
  participant_id <- strsplit(file, "_")[[1]][1]
  # Get the name of the stimuli file
  stimuli_file <- paste0(participant_id, "_stimuli.csv")
  # Read the CSV files
  data_response <- read.csv(file, header = FALSE)
  data_stimuli <- read.csv(stimuli_file, header = FALSE)
  
  # the data was saved cumulative in each step, so we only take the last rows
  if(nrow(data_response) < 21){
    data_response <- tail(data_response, n=3)
    data_stimuli <- tail(data_stimuli, n=3)
  } else {
    data_response <- tail(data_response, n=6)
    data_stimuli <- tail(data_stimuli, n=6)
  }
  
  # Convert the data into a vector
  response <- as.vector(t(data_response))
  stimuli <- as.vector(t(data_stimuli))
 
  # Create Block column
  block <- c()
  for (i in 1:nrow(data_response)){
    curr_block <- c(rep(i, ncol(data_response[i,])))
    block <- c(block, curr_block)
  }
  
  # Combine everything into one data frame
  dummy <- data.frame(response)
  dummy$stimuli <- stimuli
  dummy$participant_id <- participant_id
  dummy$block <- block
  
  # Append the data to combined_data
  localization_data <- rbind(localization_data, dummy)
}

# Sort ascending by participantID
localization_data$participant_id <- as.numeric(localization_data$participant_id)
localization_data <- localization_data[order(localization_data$participant_id), ]

# Save the combined data frame as a CSV file
#write.csv(localization_data, "localization_combined.csv", row.names = FALSE)
#localization_data
```

# Summary statistics

Mean accuracies.

```{r}
# Mean accuracy all trials
# Compute the number of lines where "response" and "stimuli" are equal
num_equal <- sum(localization_data$response == localization_data$stimuli)
# Compute the total number of lines
num_total <- nrow(localization_data)
# Compute the percentage of lines where "response" and "stimuli" are equal
percent_equal <- (num_equal / num_total) * 100
# Print the result
cat(sprintf("The accuracy over all trials is %.2f%%.", percent_equal))

# Mean accuracy all trials per participant
# Compute the percentage of lines where "response" and "stimuli" are equal for each participant
accuracy <- aggregate(localization_data$response == localization_data$stimuli, by = list(localization_data$participant_id), mean)
# Rename the columns of the result
colnames(accuracy) <- c("participant_id", "accuracy")
# Print the result
cat("\n\n")
summary(accuracy$accuracy)

# Boxplot mean accuracy
ggplot(accuracy, aes(y = accuracy)) +
  geom_boxplot() +
  labs(y = "Mean accuracy") +
  ggtitle("Boxplot of mean accuracies")
```

Calculate and visualize accuracies for first three blocks per participant.

```{r}
# Compute the maximum block number for each participant
max_blocks <- aggregate(localization_data$block, by = list(localization_data$participant_id), max)
# Compute the percentage of rows for each participant where stimuli and response are equal in the first 3 blocks
accuracy_first3 <- NULL
for (i in 1:nrow(max_blocks)) {
  curr_participant_id <- max_blocks[i, "Group.1"]
  selected_rows <- subset(localization_data, participant_id == curr_participant_id & block <= 3)
  equal_rows_count <- sum(selected_rows$response == selected_rows$stimuli)
  total_rows_count <- nrow(selected_rows)
  equal_rows_percent <- equal_rows_count / total_rows_count * 100
  accuracy_first3 <- rbind(accuracy_first3, data.frame(participant_id = curr_participant_id, accuracy = equal_rows_percent))
}

# Create a new variable to mark participant with accuracy below 90
accuracy_first3$color <- ifelse(accuracy_first3$accuracy < 90, "red", "default")

# Visualize the accuracies of the first 3 blocks
ggplot(accuracy_first3, aes(x = as.factor(participant_id), y = accuracy, fill = color)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("red" = "#F8766D", "default" = "#545454")) +
  labs(title = "Mean accuracy in first 3 trials by participant", x = "Participant ID", y = "Mean accuracy in percent") +
  coord_cartesian(ylim = c(70,100)) +
  geom_hline(yintercept = 90, color = "red", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(fill = FALSE)
```

Calculate and visualize accuracies for last three blocks per participant.

```{r}
# Compute the percentage of rows for each participant where stimuli and response are equal in the last 3 blocks
accuracy_last3 <- NULL
for (i in 1:nrow(max_blocks)) {
  curr_participant_id <- max_blocks[i, "Group.1"]
  max_block <- max_blocks[i, "x"]
  selected_rows <- subset(localization_data, participant_id == curr_participant_id & block >= (max_block - 2))
  equal_rows_count <- sum(selected_rows$response == selected_rows$stimuli)
  total_rows_count <- nrow(selected_rows)
  equal_rows_percent <- (equal_rows_count / total_rows_count) * 100
  accuracy_last3 <- rbind(accuracy_last3, data.frame(participant_id = curr_participant_id, accuracy = equal_rows_percent))
}

# Create a new variable to mark participant with accuracy below 90
accuracy_last3$color <- ifelse(accuracy_last3$accuracy < 90, "red", "default")

# Visualize the accuracies of the first 3 blocks
ggplot(accuracy_last3, aes(x = as.factor(participant_id), y = accuracy, fill = color)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("red" = "#F8766D", "default" = "#545454")) +
  labs(title = "Mean accuracy in last 3 trials by participant", x = "Participant ID", y = "Mean accuracy in percent") +
  coord_cartesian(ylim = c(85,100)) +
  geom_hline(yintercept = 90, color = "red", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(fill = FALSE)
```

Create a confusion matrix with counts and probabilities.

```{r}
# Create a table of counts for each combination of stimuli and response
count_table <- table(localization_data$stimuli, localization_data$response)

# Put "no response" column last
count_table <- count_table[,c(1,2,4,5,3)]

# Print the result
print(count_table)
```
```{r}
# Convert the table into a data frame for ggplot2
count_df <- as.data.frame(melt(count_table))

# Convert Var1 and Var2 to character
count_df$Var1 <- as.character(count_df$Var1)
count_df$Var2 <- as.character(count_df$Var2)

# Reorder the factor levels in Var2
count_df$Var2 <- factor(count_df$Var2, levels = c("down", "left", "right", "up", "no response"))

# Add a new column specifying whether or not each cell is on the diagonal
# Exclude "no response" column from the diagonal
count_df$on_diagonal <- count_df$Var1 == count_df$Var2 & count_df$Var2 != "no response"

# Create the plot
ggplot(count_df, aes(Var2, Var1, fill = on_diagonal)) +
  geom_tile(aes(fill = on_diagonal)) +
  scale_fill_manual(values = c("steelblue", "orange"), guide = FALSE) +
  geom_text(aes(label = round(value, 3)), size = 4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Predicted", y = "Actual") +
  ggtitle("Confusion Matrix (counts)")
```

```{r}
# Calculate how often 2 locations got confused.
up_right <- count_table["up","right"] + count_table["right","up"]
up_down <- count_table["up","down"] + count_table["down","up"]
up_left <- count_table["up","left"] + count_table["left","up"]
right_down <- count_table["right","down"] + count_table["down","right"]
right_left <- count_table["right","left"] + count_table["left","right"]
down_left <- count_table["down","left"] + count_table["left","down"]
noresponse <- sum(count_table[, "no response"])
right <- count_table["right", "right"]
left <- count_table["left", "left"]

cat(sprintf("Up and Right got confused %i times. \nUp and Down got confused %i times. \nUp and Left got confused %i times. \nRight and Down got confused %i times. \nRight and Left got confused %i times. \nDown and Left got confused %i times. \nIn %i trials the participant answered with no response. \nThe Right vibration motor was identified correctly most frequently, namely %i times. \nOverall participants correclty identified the Left motor the least. It was identified correctly %i times.", 
            up_right, up_down, up_left, right_down, right_left, down_left, noresponse, right, left))
```

Now for probabilities.

```{r}
# Create a confusion matrix by dividing the count table by the row sums
confusion_matrix <- prop.table(count_table, 1) * 100

# Print the result
confusion_matrix
```

```{r}
# Convert the matrix into a data frame for ggplot2
confusion_df <- as.data.frame(melt(confusion_matrix))

# Convert Var1 and Var2 to character
confusion_df$Var1 <- as.character(confusion_df$Var1)
confusion_df$Var2 <- as.character(confusion_df$Var2)

# Reorder the factor levels in Var2
confusion_df$Var2 <- factor(confusion_df$Var2, levels = c("down", "left", "right", "up", "no response"))

# Add a new column specifying whether or not each cell is on the diagonal
# Exclude "no response" column from the diagonal
confusion_df$on_diagonal <- confusion_df$Var1 == confusion_df$Var2 & confusion_df$Var2 != "no response"

# Create the heatmap
ggplot(confusion_df, aes(Var2, Var1, fill = on_diagonal)) +
  geom_tile(aes(fill = on_diagonal)) +
  scale_fill_manual(values = c("steelblue", "orange"), guide = FALSE) +
  geom_text(aes(label = round(value, 3)), size = 4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Predicted", y = "Actual") +
  ggtitle("Confusion Matrix (percentages)")
```

```{r}
# Calculate how often 2 locations got confused.
up_right <- confusion_matrix["up","right"] + confusion_matrix["right","up"]
up_down <- confusion_matrix["up","down"] + confusion_matrix["down","up"]
up_left <- confusion_matrix["up","left"] + confusion_matrix["left","up"]
right_down <- confusion_matrix["right","down"] + confusion_matrix["down","right"]
right_left <- confusion_matrix["right","left"] + confusion_matrix["left","right"]
down_left <- confusion_matrix["down","left"] + confusion_matrix["left","down"]
noresponse <- sum(confusion_matrix[, "no response"])
right <- confusion_matrix["right", "right"]
left <- confusion_matrix["left", "left"]

cat(sprintf("Up and Right got confused ~%.2f%% of the time. \nUp and Down got confused ~%.2f%% of the time. \nUp and Left got confused ~%.2f%% of the time. \nRight and Down got confused ~%.2f%% of the time. \nRight and Left got confused ~%.2f%% of the time. \nDown and Left got confused ~%.2f%% of the time. \nIn ~%.2f%% of trials the participant answered with no response. \nThe Right vibration motor was identified with the highest accuracy of ~%.2f%%. \nOverall participants identified the Left motor with the lowest acccuracy of ~%.2f%%.", up_right, up_down, up_left, right_down, right_left, down_left, noresponse, right, left))
```

Connecting localization and grasping performances.

```{r}
# filter for only successes in tactile condition
mixed_data <- grasping_data %>% filter(success=="success", condition=="tactile") %>% group_by(participant_id) %>% summarize(mean_time = mean(time)*1000)
mixed_data$acc_first3 <- accuracy_first3$accuracy
mixed_data$acc_last3 <- accuracy_last3$accuracy
mixed_data$only3 <- ifelse(mixed_data$acc_first3 == mixed_data$acc_last3, "yes", "no")

# scatterplot: y = grasping_RTs, x = mean acc (first 3 blocks)
mixed_data %>%
  ggplot(aes(x = acc_first3, y = mean_time, color=only3)) +
  geom_point() +
  labs(
    title = "Mean trial times per mean localization accuracy (first 3 blocks)", 
    x = "\n Mean accuracy (%)", 
    y = "Mean trial times (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = round(mixed_data$acc_first3, 1)) +
  scale_color_discrete(labels = c("6 localization blocks", "3 localization blocks"))

# scatterplot: y = grasping_RTs, x = mean acc (last 3 blocks)
mixed_data %>%
  ggplot(aes(x = acc_last3, y = mean_time, color=only3)) +
  geom_point() +
  labs(
    title = "Mean trial times per mean localization accuracy (last 3 blocks)", 
    x = "\n Mean accuracy (%)", 
    y = "Mean trial times (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = round(mixed_data$acc_last3, 1)) +
  scale_color_discrete(labels = c("6 localization blocks", "3 localization blocks"))
```

```{r}
# barplot: y = mean RT, x = mean accuracy (last 3 blocks)
mixed_data %>% group_by(acc_last3) %>% summarize(mean_time = mean(mean_time))
mixed_data %>% group_by(acc_last3) %>% summarize(mean_time = mean(mean_time)) %>%
  ggplot(aes(x = acc_last3, y = mean_time)) +
  geom_bar(stat = "identity", position = "dodge", colour="#CC9900", fill="#FFCC66") +
  labs(
    title = "Mean trial times per mean localization accuracy", 
    x = "Mean accuracy (%)", 
    y = "Mean trial times (ms) \n",
    ) +
  theme_linedraw() +
  scale_x_continuous(breaks = round(mixed_data$acc_last3, 1)) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

```{r}
# have a look at the errors per accuracy
fails_df <- grasping_data %>% 
  select(-1) %>% 
  filter(success=="fail", condition=="tactile") %>% 
  group_by(participant_id) %>% 
  count() %>% 
  rename_at('n', ~'fails') %>% 
  rbind(data.frame(participant_id = c(23,29), fails = c(0,0))) %>% 
  arrange(participant_id)

# add to mixed data
mixed_data$fails <- fails_df$fails

# participants scatterplot: y = mean fails, x = mean acc (last 3 blocks)
mixed_data %>%
  ggplot(aes(x = acc_last3, y = fails, colour=only3)) +
  geom_point() +
  labs(
    title = "Mean fails per mean localization accuracy", 
    x = "\n Mean accuracy (%)", 
    y = "Mean number of fails \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = round(mixed_data$acc_last3, 1)) +
  scale_y_continuous(breaks = c(0:15)) +
  scale_color_discrete(labels = c("6 localization blocks", "3 localization blocks"))
```

```{r}
# barplot: y = mean fails, x = mean acc
mixed_data %>% group_by(acc_last3) %>% summarize(mean_fails = mean(fails))
mixed_data %>% group_by(acc_last3) %>% summarize(mean_fails = mean(fails)) %>% 
  ggplot(aes(x = acc_last3, y = mean_fails)) +
  geom_bar(stat = "identity", position = "dodge", colour="#CC9900", fill="#FFCC66") +
  labs(
    title = "Mean fails per mean localization accuracy", 
    x = "Mean accuracy (%)", 
    y = "Mean number of fails \n",
    ) +
  theme_linedraw() +
  scale_x_continuous(breaks = round(mixed_data$acc_last3, 1)) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

```{r}
# idea: create a scatterplot in 3D to observe correlation between errors in both tasks in form of clusters -- if participants failed in correctly identifying "left" direction in the localization task, do they also make more mistakes regarding position 4 in the grasping task?

# 1. Create grasping error counts per object location
# rename locations
pos_data <- grasping_data # dummy
pos_data$location[pos_data$location == "rep_1"] = "1"
pos_data$location[pos_data$location == "rep_2"] = "2"
pos_data$location[pos_data$location == "rep_3"] = "3"
pos_data$location[pos_data$location == "rep_4"] = "4"
pos_data$location[pos_data$location == "rep_5"] = "5"
pos_data$location[pos_data$location == "rep_6"] = "6"
pos_data$location[pos_data$location == "rep_7"] = "7"
pos_data$location[pos_data$location == "rep_8"] = "8"
pos_data$location[pos_data$location == "rep_9"] = "9"

# create a failed column (necessary for 0 counts in the code below)
pos_data$failed <- ifelse(pos_data$success == "fail", 1, 0)

# grasping error counts per participant per object location (left, right, up, down)
pos_data <- pos_data %>% filter(condition=="tactile") %>% complete(participant_id, location, fill = list(failed = 0)) %>% group_by(participant_id, location) %>% summarise(grasping_errors = sum(failed, na.rm = TRUE)) %>% filter(location == 2 | location == 4 | location == 6 | location == 8)

# rename locations to localization task directions for merging
pos_data$location[pos_data$location == 2] = "up"
pos_data$location[pos_data$location == 4] = "left"
pos_data$location[pos_data$location == 6] = "right"
pos_data$location[pos_data$location == 8] = "down"


# 2. Create localization error counts
# only last 3 blocks for all participants are taken into account
loc_data <- localization_data %>% filter(!(participant_id==5 & (block==1 | block==2 | block==3)),
                             !(participant_id==10 & (block==1 | block==2 | block==3)),
                             !(participant_id==11 & (block==1 | block==2 | block==3)),
                             !(participant_id==24 & (block==1 | block==2 | block==3)))

# determine failed directions (location) in localization trials 
loc_data$location <- case_when(
  (loc_data$stimuli == "left" & loc_data$response != "left") ~ "left",
  (loc_data$stimuli == "right" & loc_data$response != "right") ~ "right",
  (loc_data$stimuli == "up" & loc_data$response != "up") ~ "up",
  (loc_data$stimuli == "down" & loc_data$response != "down") ~ "down",
  TRUE ~ "none")

# create a failed column (necessary for 0 counts in the code below)
loc_data$failed <- ifelse(loc_data$response == loc_data$stimuli, 0, 1)

# localization error counts per participant per command direction (left, right, up, down)
loc_data <- loc_data %>% filter(failed==1) %>% complete(participant_id, location, fill = list(failed = 0)) %>% group_by(participant_id,location) %>% summarise(localization_errors = sum(failed, na.rm = TRUE))


# 3. Merge grasping and localization errors per participant and location/direction
cluster_data <- merge(pos_data, loc_data, by = c("participant_id", "location"), all = TRUE)
cluster_data[is.na(cluster_data)] <- 0
cluster_data <- cluster_data %>% filter(!(grasping_errors == 0 & localization_errors == 0))
cluster_data
```

```{r}
# boxplot correlation coefficients between localization and grasping errors
# calculation: graspingERR / localizationERR (transformed 0's to -1's) --> values close to 1 are positively correlating.
coeffs <- cluster_data %>% mutate_all(~ ifelse(. == 0, -1, .)) %>% select(-participant_id) %>% group_by(location) %>% summarise(coeff = grasping_errors / localization_errors)

coeffs %>% 
  ggplot(aes(x = location, y = coeff)) +
  geom_boxplot() +
  labs(x = "Localization direction / Grasping location", y = "Correlation") +
  ggtitle("Mean positive correlation between localization and grasping errors")

cor(cluster_data$grasping_errors, cluster_data$localization_errors)
```
Interpretation: The plot shows that the errors made in both tasks seem to be unrelated. The correlation between both error types is -0.308, slightly negatively correlated, which means that for errors made in one task the error count in the other task is going into the other direction. This might mean that participants in the localization task efficiently learned to identify vibration directions (0 was swapped with -1 values for calculation, so overall negative correlation values mean overall low error counts).

```{r}
# 3D scatterplot: x = command direction / object location, y = grasping errors, z = localization errors
cluster_data %>% 
  plot_ly(x = ~location, y = ~grasping_errors, z = ~localization_errors, type = "scatter3d", 
          color = ~location, mode = "markers", marker = list(size = 4)) %>% 
  layout(
    scene = list(
      xaxis = list(title = "Location / Direction", tickvals = unique(cluster_data$location)),
      yaxis = list(title = "Grasping errors", tickvals = unique(cluster_data$grasping_errors)),
      zaxis = list(title = "Localization errors", tickvals = c(1,2,3)),
      showaxes = FALSE,
      showbackground = FALSE  # Turn off the background color
    ),
    title = "\n Correlation scatter plot of errors in tasks"
  )

# data points are overlaying each other
```

