---
title: "OptiVisT questionnaire analysis"
author: "Florian Pätzold, Ramon Zacharias"
date: "2023-09-05"
output:
  html_document: default
  pdf_document: default
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Import all necessary packages.

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(psych)
library(corrr)
library(ggcorrplot)
library(FactoMineR)
```

Load the data.

```{r}
# Set working directory to folder with localization files
SAVE <- paste0(getwd(), "/Plots/")
setwd("/Users/florian/Documents/Studium/NBP/Projects/OptiVisT/FloBox/Experiment/Data/Questionnaire/")
# Read in the data
questionnaire_data <- read.csv("questionnaire.csv")
```

# Summary statistics

We get summary stats for each question and plot them in a box plot. Some additional comments: for analysis we would probably rather take the median instead of the mean into account, as the mean as central tendency measure might not capture the data's characteristics, e.g. when there are many 1s and 5s, the mean is 3, thus the answer seems to be neutral. The median in fact does not solve that, but might give a better picture (Sullivan & Artino, 2014). Furthermore, Likert scale (ordinal) data usually cannot be seen as interval data, but parametric tests can be used with sufficient sample size as they have higher power (Norman, 2010; de Winter & Dodou, 2010).
However, if mean and median do not differ by much, we get the confirmation that the median captures the data's characteristics as well. In our data, this is the case.

```{r}
# Define a custom summary function that calculates statistics with NA values removed
custom_summary <- function(x) c(mean = mean(x, na.rm = TRUE), sd = sd(x, na.rm = TRUE), median = median(x, na.rm = TRUE), min = min(x, na.rm = TRUE), max = max(x, na.rm = TRUE))

# Use sapply with the custom summary function
summary_stats <- sapply(questionnaire_data[, 2:18], custom_summary)

# Convert the result to a data frame for a cleaner display
summary_df <- as.data.frame(summary_stats)
summary_df
```

```{r}
# Further get the 1st and 3rd quartile together with mean, median min, max stats
summary(questionnaire_data[,2:18]) # exclude two last participants, as they are blind Ps
```

```{r}
# Plot each question with box plot
for (col in colnames(questionnaire_data)[2:18]) {
  plot_data <- data.frame(Response = questionnaire_data[, col], Question = gsub("\\.", " ", col))
  plot_title <- gsub("^X", "", col)
  plot_title <- gsub("\\.", " ", plot_title)

  plot <- ggplot(plot_data, aes(x = factor(1), y = Response)) +
    geom_boxplot(fill = "lightblue", color = "black", outlier.shape = NA) +
    geom_jitter(width = 0.5, height = 0, size = 3, alpha = 0.6, color = "red") +  # add points
    labs(title = plot_title, x = "", y = "Response") +
    #theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    ylim(1, 5)  # set y-axis limits to 1 and 5

  print(plot)
}
```

# Open-ended questions

We here qualitatively analyse the comments made in the open-ended questions, regarding the experience with the tactile bracelet and the grasping task.

```{r}
print(questionnaire_data$`Please.describe.your.general.experience.with.the.tactile.bracelet.itself..`)
```

*Summary:*  
- some participants had troubles to distinguish vibration directions especially after fatigue (-> blending)  
- differences in wrist circumferences can lead to perception of too large/small motors and/or vibration intensity and distinguishing directions can again be influenced by that  
- grasping cue -> some participants started moving up first --> change it to something else (e.g. all motors vibrate, last direction pulses)  
- some participants felt too much vibration intensity -> numbs wrist (and forearm) -> fatigues -> harder to distinguish -> less intuitive  
- one P with troubles distinguishing spoken left-right reported it was easier with the bracelet  

```{r}
print(questionnaire_data$`Please.describe.your.general.experience.with.the.grasping.task`)
```

**Summary:**  
- grasping anticipation because of regular pattern  
- one P suggests varying vibration intensity based off of distance to target (depth)  
- unnatural head movements during the task needed some attention (because of narrow camera angle)  
- participants easily developed a spatial mapping of the shelf and arrangement of fruits  

```{r}
print(questionnaire_data$`Do.you.have.any.other.comments.for.us.`)
```

# Questionnaire constructs

We perform latent variable inference via FA and PCA to extract underlying constructs from numerical responses to the questionnaire items.

## Defining intuitive constructs

Depending on the research goals, we are defining the intuitive constructs of the questions for comparison with the analysis results.

1. **Bracelet usability**
- I felt confident using the tactile bracelet to locate and grasp an object.
- I relied on the vibration signals from the tactile bracelet to grasp an object.
- Mainly my intuition guided my hand movement during the task.
- I could identify the vibration locations without much effort.
- It was difficult for me to interpret the vibration cues from the tactile bracelet.

2. **Bracelet design** (Concept & Comfort)
- The tactile bracelet is comfortable to wear.
- The vibration from tactile bracelet felt uncomfortable on my hand.
- I could feel the vibration cues on my wrist from the tactile bracelet.
- The intensity of vibration varied strongly.
- I found the vibration intensity to be consistent.

3. **Experiment design**
- I was in fear of hitting the shelf or objects during the task.
- I felt secure during the task.
- The practice trials were sufficient to get comfortable with using the tactile feedback.
- I was able to develop a spatial understanding of the shelf.
- I waited for instructions before reaching to grasp an object.
- I was anticipating the grasping motion.
- The vibration motors were positioned correctly on my arm.

## Factor Analysis

```{r}
# remove NAs
data <- na.omit(questionnaire_data[,2:18])
```

Conduct factor analysis to explore the underlying structure of the questionnaire. This will help confirm or refine the identified constructs.

```{r, eval = FALSE, include = FALSE}
fa_data <- na.omit(questionnaire_data)[2:18]
rownames(fa_data) <- NULL

determine_factor <- function(data, data_col){
  ### Check an FA item for its factor by changing coding direction of the item. ###
  
  # sanity check
  sanity <- data
  col = sanity[data_col][,1]
  
  # 
  for (i in seq_along(col)) {
    val <- col[i]
    ifelse(val == 1, col[i] <- 5, 
           ifelse(val == 2, col[i] <- 4, 
                  ifelse(val == 4, col[i] <- 2, 
                         ifelse(val == 5, col[i] <- 1, col[i] <- 3))))
  }
  sanity[data_col] <- col
  
  # FA
  loadings_original <- fa(data, nfactors = 4, rotate = "varimax", fm = "gls", max.iter = 1000)
  loadings_sanity <- fa(sanity, nfactors = 4, rotate = "varimax", fm = "gls", max.iter = 1000)
  print(loadings_original)
  print(loadings_sanity)
}

# now do that for each question / item
for (name in colnames(fa_data)) {
  #print(name)
  determine_factor(fa_data, name)
}
```

```{r}
# remove indices
rownames(data) <- NULL

fa_model <- fa(data, nfactors = 3, rotate = "varimax", fm = "pa", max.iter = 1000)
fa_model
```
*Interpretation:* We choose to select the questions of each factor with correlation equal or above 0.2 (at least slight positive correlation) and therefore come up with the following groupings:

**PA1: user confidence**  
- I.could.identify.the.vibration.locations.without.much.effort. 0.82  
- X.I.felt.confident.using.the.tactile.bracelet.to.locate.and.grasp.an.object. 0.64  
- The.practice.trials.were.sufficient.to.get.comfortable.with.using.the.tactile.feedback. 0.61  
- I.waited.for.instructions.before.reaching.to.grasp.an.object. 0.29  
- I.was.in.fear.of.hitting.the.shelf.or.objects.during.the.task. 0.25  

**PA2: comfort**  
- The.tactile.bracelet.is.comfortable.to.wear. 0.90  
- I.found.the.vibration.intensity.to.be.consistent. 0.50  
- I.relied.on.the.vibration.signals.from.the.tactile.bracelet.to.grasp.an.object. 0.37  
- X.I.felt.confident.using.the.tactile.bracelet.to.locate.and.grasp.an.object. 0.35  
- The.vibration.motors.were.positioned.correctly.on.my.arm. 0.27  
- I.waited.for.instructions.before.reaching.to.grasp.an.object. 0.25  


**PA3: learning / intuition**  
- Mainly.my.intuition.guided.my.hand.movement.during.the.task. 0.74  
- I.was.able.to.develop.a.spatial.understanding.of.the.shelf. 0.74  
- I.was.anticipating.the.grasping.motion. 0.30  
- I.was.in.fear.of.hitting.the.shelf.or.objects.during.the.task. 0.21  
- I.felt.secure.during.the.task. 0.20  


## Principal Component Analysis

We perform PCA (dimensionality reduction) to extract the PCs that explain more than half the variance in the data, possible more. 

```{r}
# function for displaying each PC and its ordered correlation coefficients
# returns the correlation matrices after PCA and reconstructed
extract_PC <- function(data, n_comps, thres=-1){
  # normalize data and perform PCA
  corr_matrix <- cor(scale(data))
  data.pca <- princomp(corr_matrix)
  print(summary(data.pca))
  # extract correlation loadings
  loadings <- data.pca$loadings[, 0:n_comps]
  
  # output each PC
  for (i in 1:n_comps) {
    print(data.frame(item=loadings[,0], PC=loadings[,i]) %>% arrange(desc(PC)) %>% filter(PC >= thres))
  }
  
  # use different PCA method for data reconstruction
  res <- prcomp(data, scale. = TRUE)
  corr_matrix_reconstructed <- cor(res$x[, 0:n_comps] %*% t(res$rotation[, 0:n_comps]))
  
  return(list("pca" = res, "loadings" = loadings, "normal" = corr_matrix, "reconstructed" = corr_matrix_reconstructed))
}

# extract 3 PCs
res <- extract_PC(data, 3)
```

```{r}
eigenvalues <- res$pca$sdev^2
#retained_components <- sum(eigenvalues > 1)
eigenvalues

plot(eigenvalues, type = "b", pch = 19, xlab = "Principal Component", ylab = "Eigenvalue", main = "Scree Plot")
abline(h = 1, col = "red", lty = 2)
abline(h = 2, col = "green", lty = 2)
print(paste0("Number of PCs with EV > 1: ", sum(eigenvalues > 1)))
print(paste0("Number of PCs with EV > 2: ", sum(eigenvalues > 2)))
```

In this case, 2 PCs explain ~57%, 3 PCs ~70% of the variance in the data. The screeplot shows cut-off lines for eigenvalues > 1 and 2, following Kaiser's stopping rule. 7 PCs lie above Kaiser's EV of 1, explaining ~95% of the variance. We here choose 3 PCs, as these are the ones above EV of 2, and it lines up with our a priori selection for the questionnaire design. This indicates that the questions were not formulated in a way such that a single question exclusively investigates one of the intuitive groupings but as 3 PCs explain more than 70% of cumulative variance and afterwards explained-variance differences are lower than 10%, we confidently stick to 3 PCs.

```{r}
# shorten question names
qs <- c("hit_shelf", "secure", "comfortable", "vib_uncomfortable", "motor_pos", "feel_vib", "vib_intensity", "intensity_cons", "vib_locs", "vib_interp", "practice", "reliance", "intuition", "confidence", "spatial", "instr_wait", "anticipation")

#qs <- c("fear_of_hitting_shelf", "sense_of_task_security", "bracelet_comfort", "vibration_discomfort", "correct_motor_positioning", "ability_to_feel_vibration", "vibration_intensity_variance", "vibration_intensity_consistency", "vibration_location_identifiability", "vibration_interpretability", "practice_sufficiency", "reliance_on_vibrations", "guidance_by_intuition", "bracelet_usage_confidence", "development_of_spatial_map", "instruction_patience", "grasping_anticipation")

pcs <- data.frame(Question = rownames(res$loadings), PC1=res$loadings[,1], PC2=res$loadings[,2], PC3=res$loadings[,3])
# convert to long format
pcs <- pcs %>% pivot_longer(cols = starts_with("PC"), names_to = "PC", values_to = "Value")
pcs$Question <- factor(pcs$Question, levels = rownames(res$loadings))

# Create a faceted barplot
loads <- ggplot(pcs, aes(x = Question, y = Value, fill = PC)) +
  geom_bar(stat = "identity", width = 0.7) +
  labs(#title = "Component loadings of first 3 PCs",
       x = NULL,
       y = "Correlation coefficient value") +
  facet_wrap(~ PC, scales = "free_y", ncol = 1) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_x_discrete(labels = qs)
loads

#ggsave(paste0(SAVE,"comp_loadings.jpeg"), plot = loads, dpi = 600)
```


We then plot the correlation matrices after PCA, of the reconstructed data and re-ordered (using hierarchical clustering) to visualize the PCs (Kording et al., 2018).

```{r}
# comfortable, intensity_cons, feel_vib, motor_pos, reliance, confidence, instr_wait

# plot corr matrix after pca
pca_before <- ggcorrplot(res$normal) + 
  scale_x_discrete(labels = qs) +
  scale_y_discrete(labels = qs)
pca_before

# plot reconstructed corr matrix
pca_after <- ggcorrplot(res$reconstructed) + 
  scale_x_discrete(labels = qs) +
  scale_y_discrete(labels = qs)
pca_after

# Perform hierarchical clustering and get order
#hclust_result <- hclust(dist(res$reconstructed))
#order <- hclust_result$order
order <- c(10,17,4,7,6,3,8,5,12,14,16,1,2,13,15,9,11) # re-order to visualize the 3 PCs
corr_reordered <- res$reconstructed[order,order]

# plot re-ordered reconstructed corr matrix
reordered <- ggcorrplot(corr_reordered) + 
  scale_x_discrete(labels = qs[order]) +
  scale_y_discrete(labels = qs[order])
reordered

#ggsave(paste0(SAVE,"corr_mat_reordered.jpeg"), plot = reordered, dpi = 600)
```

### Interpretation

**Question keys**  
- 1 hit_shelf: I.was.in.fear.of.hitting.the.shelf.or.objects.during.the.task.  
- 2 secure: I.felt.secure.during.the.task.  
- 3 comfortable: The.tactile.bracelet.is.comfortable.to.wear.  
- 4 vib_uncomfortable: The.vibration.from.tactile.bracelet.felt.uncomfortable.on.my.hand.  
- 5 motor_pos: The.vibration.motors.were.positioned.correctly.on.my.arm.  
- 6 feel_vib: I.could.feel.the.vibration.cues.on.my.wrist.from.the.tactile.bracelet.  
- 7 vib_intensity: The.intensity.of.vibration.varied.strongly.  
- 8 intensity_cons: I.found.the.vibration.intensity.to.be.consistent.  
- 9 vib_locs: I.could.identify.the.vibration.locations.without.much.effort.  
- 10 vib_interp: It.was.difficult.for.me.to.interpret.the.vibration.cues.from.the.tactile.bracelet.  
- 11 practice: The.practice.trials.were.sufficient.to.get.comfortable.with.using.the.tactile.feedback.  
- 12 reliance: I.relied.on.the.vibration.signals.from.the.tactile.bracelet.to.grasp.an.object.  
- 13 intuition: Mainly.my.intuition.guided.my.hand.movement.during.the.task.  
- 14 confidence: X.I.felt.confident.using.the.tactile.bracelet.to.locate.and.grasp.an.object.  
- 15 spatial: I.was.able.to.develop.a.spatial.understanding.of.the.shelf.  
- 16 instr_wait: I.waited.for.instructions.before.reaching.to.grasp.an.object.  
- 17 anticipation: I.was.anticipating.the.grasping.motion.  

**PC1: bracelet usability**  
- It.was.difficult.for.me.to.interpret.the.vibration.cues.from.the.tactile.bracelet. 0.38
- The.vibration.from.tactile.bracelet.felt.uncomfortable.on.my.hand. 0.30  
- I.was.anticipating.the.grasping.motion. 0.21  
- The.intensity.of.vibration.varied.strongly. 0.16  

**PC2: task & bracelet confidence**  
- The.vibration.from.tactile.bracelet.felt.uncomfortable.on.my.hand. 0.36
- The.intensity.of.vibration.varied.strongly. 0.34  
- I.could.identify.the.vibration.locations.without.much.effort. 0.29  
- I.was.in.fear.of.hitting.the.shelf.or.objects.during.the.task. 0.28  
- The.practice.trials.were.sufficient.to.get.comfortable.with.using.the.tactile.feedback. 0.16  

**PC3: learning intuition**  
- Mainly.my.intuition.guided.my.hand.movement.during.the.task. 0.55
- I.was.able.to.develop.a.spatial.understanding.of.the.shelf. 0.48  
- I.was.in.fear.of.hitting.the.shelf.or.objects.during.the.task. 0.37  
- I.could.identify.the.vibration.locations.without.much.effort. 0.29  
- The.practice.trials.were.sufficient.to.get.comfortable.with.using.the.tactile.feedback. 0.16  
- I.felt.secure.during.the.task. 0.14  


### Visualize PCs

```{r}
# https://www.statology.org/principal-components-analysis-in-r/

# bi-plot
res$pca$x <- -1*res$pca$x
biplot(res$pca, scale = 0, ylabs = qs) # 1 vs 2
biplot(res$pca, choices = c(1, 3), scale = 0, ylabs=qs) # 1 vs 3
biplot(res$pca, choices = c(2, 3), scale = 0, ylabs=qs) # 2 vs 3
```
# Summary statistics

## Data preparation

We clean the data and wrangle it into another format

```{r}
clean_data <- data.frame()
for (col in colnames(questionnaire_data)[2:18]) {
  questions <- data.frame(Response = questionnaire_data[, col], Question = gsub("\\.", " ", col))
  title <- gsub("^X", "", col)
  title <- gsub("\\.", " ", title)
  clean_data <- rbind(clean_data, questions[-1,])
}
rownames(clean_data) <- NULL

# rename one question (remove preceding 'X')
clean_data[clean_data$Question == "X I felt confident using the tactile bracelet to locate and grasp an object ",]$Question <- "I felt confident using the tactile bracelet to locate and grasp an object "
```

Add underlying constructs.

```{r}
coded_data <- clean_data

# PC1: tactile signals, usability, interpretability
usability <- c(
  "The intensity of vibration varied strongly ",
  "The vibration from tactile bracelet felt uncomfortable on my hand ",
  "I was anticipating the grasping motion ",
  "It was difficult for me to interpret the vibration cues from the tactile bracelet "  
)

# PC2: experimental design, bracelet design, consistency
task <- c(
  "I could feel the vibration cues on my wrist from the tactile bracelet ",
  "The tactile bracelet is comfortable to wear ",
  "I found the vibration intensity to be consistent ",
  "The vibration motors were positioned correctly on my arm ",
  "I relied on the vibration signals from the tactile bracelet to grasp an object ",
  "I felt confident using the tactile bracelet to locate and grasp an object ",
  "I waited for instructions before reaching to grasp an object "
)

# PC3: security, confidence, practice
confidence <- c(
  "I was in fear of hitting the shelf or objects during the task ",
  "I felt secure during the task ",
  "Mainly my intuition guided my hand movement during the task ",
  "I was able to develop a spatial understanding of the shelf ",
  "I could identify the vibration locations without much effort ",
  "The practice trials were sufficient to get comfortable with using the tactile feedback "
)

coded_data$PC1 <- coded_data$Question %in% usability
coded_data$PC2 <- coded_data$Question %in% task
coded_data$PC3 <- coded_data$Question %in% confidence
```

### Coding direction

For negative questions we are changing the coding direction, such that a higher value consistently represents higher agreement with the underlying construct, e.g. "The intensity of vibration varied strongly" responses will be reversed in order. Note: Grasping anticipation during the task under the experiment design construct is a bad thing.

```{r}

conditions <- coded_data$Question %in% c(
  "It was difficult for me to interpret the vibration cues from the tactile bracelet ",
  "The vibration from tactile bracelet felt uncomfortable on my hand ",
  "The intensity of vibration varied strongly ",
  "I was in fear of hitting the shelf or objects during the task ",
  "I was anticipating the grasping motion "
)

# change coding direction
coded_data$Response[conditions] <- 
  ifelse(coded_data$Response[conditions] == 1, 5, 
         ifelse(coded_data$Response[conditions] == 2, 4, 
                ifelse(coded_data$Response[conditions] == 4, 2, 
                       ifelse(coded_data$Response[conditions] == 5, 1, 3))))

# rename questions for easier interpretability
coded_data <- coded_data %>%
  mutate(Question = case_when(
    Question == "It was difficult for me to interpret the vibration cues from the tactile bracelet " ~ "It was easy for me to interpret the vibration cues from the tactile bracelet ",
    Question == "The vibration from tactile bracelet felt uncomfortable on my hand " ~ "The vibration from tactile bracelet felt comfortable on my hand ",
    Question == "The intensity of vibration varied strongly " ~ "The intensity of vibration did not vary strongly ",
    Question == "I was in fear of hitting the shelf or objects during the task " ~ "I was not in fear of hitting the shelf or objects during the task ",
    Question == "I was anticipating the grasping motion " ~ "I was not anticipating the grasping motion ",
    TRUE ~ Question
  ))

#coded_data
```

```{r}
pc1 <- coded_data %>% filter(PC1) %>% summarize(mean = mean(Response), sd=sd(Response))
pc2 <- coded_data %>% filter(PC2) %>% summarize(mean = mean(Response), sd=sd(Response))
pc3 <- coded_data %>% filter(PC3) %>% summarize(mean = mean(Response), sd=sd(Response))

coded_data %>% summarize(mean = mean(Response), sd = sd(Response))

construct_eval <- data.frame(mean=c(pc1[,1],pc2[,1],pc3[,1]), sd=c(pc1[,2],pc2[,2],pc3[,2]), Construct=factor(c("usability", "task", "confidence"), levels = c("usability", "task", "confidence"))) %>%
  ggplot(aes(x=Construct, y = mean)) +
  geom_bar(stat = "identity", position = "dodge", colour="#CC9900", fill="#FFCC66") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width=0.2) +
  #geom_text(aes(label = round(mean, 2)), position = position_dodge(width = 1), vjust = 2, hjust=0.5) +
  labs(
    #title = "Mean participant evaluation per construct", 
    x = NULL, 
    y = "Mean response (Likert-scale) \n",
    ) +
  theme_linedraw() +
  scale_y_continuous(limits=c(0,6))

construct_eval

#ggsave(paste0(SAVE,"construct_eval.jpeg"), plot = construct_eval, dpi = 600)
```

*Interpretation:* Participants rated the usability of the bracelet lowest with a mean of 3.66 (sd = 1.32), most certainly because some participants had a hard time correctly interpreting the vibration signals as motors directly on a bone spread the signal across larger areas. Task design and the usage of the bracelet in the specific task was rated highest with 4.29 (sd = 0.88), indicating a generally good experimental design and adaptation to the bracelet. Confidence of using the bracelet was rated with 4.0 (sd = 1.21) suggesting that the idea of the bracelet and the usage itself are learnable relatively quickly. Overall, participants rated the whole experience relatively high which serves as confirmation for further developing the bracelet.

# References

- Kording, K., Blohm, G., Schrater, P., & Kay, K. (2018). Appreciating diversity of goals in computational neuroscience. OSF Preprints.  
- Norman, G. (2010). Likert scales, levels of measurement and the “laws” of statistics. Advances in health sciences education, 15(5), 625-632.  
- Sullivan, G. M., & Artino, A. R., Jr (2013). Analyzing and interpreting data from likert-type scales.   Journal of graduate medical education, 5(4), 541–542. https://doi.org/10.4300/JGME-5-4-18. 
- de Winter, J.C.F., & Dodou, D. (2010), Five-Point Likert Items: t test versus Mann-Whitney-Wilcoxon, Practical Assessment, Research and Evaluation, 15(11).  