---
title: "OptiVisT questionnaire analysis"
author: "Florian Pätzold, Ramon Zacharias"
date: "2023-09-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Import all necessary packages.

```{r}
library(ggplot2)
```

```{r}
# Set working directory to folder with localization files
setwd("/Users/florian/Documents/Studium/NBP/Projects/OptiVisT/FloBox/analysis/data/questionnaire/")
# Read in the data
questionnaire_data <- read.csv("questionnaire.csv")
```

```{r}
# Compute summary statistics for each question
summary(questionnaire_data[,2:18])
```

```{r}
# Plot each question with box plot
for (col in colnames(questionnaire_data)[2:18]) {
  plot_data <- data.frame(Response = questionnaire_data[, col], Question = gsub("\\.", " ", col))
  plot_title <- gsub("^X", "", col)
  plot_title <- gsub("\\.", " ", plot_title)

  plot <- ggplot(plot_data, aes(x = factor(1), y = Response)) +
    geom_boxplot(fill = "lightblue", color = "black", outlier.shape = NA) +
    geom_jitter(width = 0.5, height = 0, size = 3, alpha = 0.6, color = "red") +  # add points
    labs(title = plot_title, x = "", y = "Response") +
    #theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    ylim(1, 5)  # set y-axis limits to 1 and 5

  print(plot)
}
```

```{r}
print(questionnaire_data$`Please.describe.your.general.experience.with.the.tactile.bracelet.itself..`)
```
*Summary:*  
- some participants had troubles to distinguish vibration directions especially after fatigue (-> blending)  
- differences in wrist circumferences can lead to perception of too large/small motors and/or vibration intensity and distinguishing directions can again be influenced by that  
- grasping cue -> some participants started moving up first --> change it to something else (e.g. all motors vibrate, last direction pulses)  
- some participants felt too much vibration intensity -> numbs wrist (and forearm) -> fatigues -> harder to distinguish -> less intuitive  
- one P with troubles distinguishing spoken left-right reported it was easier with the bracelet  

```{r}
print(questionnaire_data$`Please.describe.your.general.experience.with.the.grasping.task`)
```
*Summary:*  
- grasping anticipation because of regular pattern  
- one P suggests varying vibration intensity based off of distance to target (depth)  
- unnatural head movements during the task needed some attention (because of narrow camera angle)  
- participants easily developed a spatial mapping of the shelf and arrangement of fruits  

```{r}
print(questionnaire_data$`Do.you.have.any.other.comments.for.us.`)
```




